### 2016-03-05 paper/abstract.tex

\begin{abstract}
\yasu{must be at least 70 and at most 150 words.}
\keywords{A; B; C.}
\end{abstract}