### 2016-03-10 paper/abstract.tex
yasu

\begin{abstract}
This paper proposes a novel MAP inference framework for Markov Random
Field (MRF) problems in parallel computing environments.  The inference
framework, dubbed Swarm Fusion method, is a natural generalization of
Fusion Move method for parallel computation. Every thread (in a case of
multi-threading environment) maintains and updates a solution. At every
iteration, each thread can generate arbitrary number of solution
proposals and take arbitrary number of concurrent solutions from the
other threads to perform multi-way fusion in updating its solution. The
framework is general, making popular existing inference techniques such
as alpha-expansion, fusion move, parallel alpha-expansion, or
hierarchical fusion, its special cases. We have evaluated the approach
against competing methods on stereo, optical flow, and layered depthmap
estimation problems. Our experiments have shown that the Swarm Fusion
method exploits the parallel computing resources and become effective
especially for challenging problems.
%\yasu{must be at least 70 and at most 150 words.}
\keywords{MRF; Fusion Move; Particle Swarm Optimization}
\end{abstract}