### 2016-03-11 paper/abstract.tex

\begin{abstract}
This paper proposes a novel MAP inference framework for Markov Random
Field (MRF) in parallel computing environments.  The inference
framework, dubbed Swarm Fusion, is a natural generalization of the
Fusion Move method. Every thread (in a case of multi-threading
environments) maintains and updates a solution. At every iteration, a
thread can generate arbitrary number of solution proposals and take
arbitrary number of concurrent solutions from the other threads to
perform multi-way fusion in updating its solution. The framework is
general, making popular existing inference techniques such as
alpha-expansion, fusion move, parallel alpha-expansion, or hierarchical
fusion, its special cases. We have evaluated the effectiveness of our
approach against competing methods on three problems of varying
difficulties, in particular, the stereo, the optical flow, and the
layered depthmap estimation problems.
%Our experiments have shown that the Swarm Fusion method outperform
%competing methods, especially for challenging problems with massive
%solution space.
% can effectively exploits the parallel computing resources and become
%effective especially for challenging problems.
%\yasu{must be at least 70 and at most 150 words.}
\keywords{MRF; Fusion Move; Particle Swarm Optimization}
\end{abstract}