### 2016-03-14 paper/results.tex
Hang

\section{Experimental results}
\label{section:results}
We have implemented the algorithms with multithreading supporting from
C++ 11 and conducted the experiments on Linux PCs with Intel Core i7
4790 processor with 4 cores. We have used the Graph-cuts optimization
code written by Veksler, using the libraries provided by Boykov and
Kolmogorov~\cite{middlebury_mrf,alpha_expansion,what_energy_can_be_min_by_gc,mrf_experimental}.
% [2] Fast Approximate Energy Minimization via Graph Cuts.
%         Y. Boykov, O. Veksler, and R. Zabih.
%         In IEEE Transactions on Pattern Analysis and Machine Intelligence
%         (PAMI), vol. 23, no. 11, pages 1222-1239, November 2001.  
%
%     [3] What Energy Functions can be Minimized via Graph Cuts?
%         V. Kolmogorov and R. Zabih. 
%         In IEEE Transactions on Pattern Analysis and Machine Intelligence
%         (PAMI), vol. 26, no. 2, pages 147-159, February 2004. 
%         An earlier version appeared in European Conference on Computer
%         Vision (ECCV), May 2002.
%
%     [4] An Experimental Comparison of Min-Cut/Max-Flow Algorithms for
%         Energy Minimization in Vision. 
%         Y. Boykov and Vladimir Kolmogorov.
%         In IEEE Transactions on Pattern Analysis and Machine Intelligence
%         (PAMI), vol. 26, no. 9, pages 1124-1137, September 2004. 
We have used the QPBO and TRW-S implementations by
Kolmogorov~\cite{QPBO, TRW-S_implementation}.

To evaluate the effectness of different parallel structures, we plot
the energy minimization process against time. We define the energy of
a parallel optimization system at a certain time as the minimum energy
of all threads at that time.

\input{result_stereo}
\input{result_optical_flow}
\input{result_layered_depthmap}