### 2016-03-14 paper/result_stereo.tex
Hang

\mysubsubsection{Stereo}

\noindent We have chosen 7 images with the resolution of $695\times555$
from the Book sequence of Middlebury stereo dataset for our
experiment~\cite{middlebury_stereo}. The number of disparity labels is
set to 256.
%
Since the order of labels is important for the expansion techniques,
we have used the same random order for all algorithms to avoid any
bias.
%
% In PAE, SF-MF, SF-SS and SF algorithms, the proposal generator in each
% thread can only generate constant-label proposals within a evenly split
% subset of all labels.

Fig.~\ref{fig:stereo_global} compares the converegence rate of the
competing methods. Note that we define the energy of a multi-threading
system to be the energy of the best solution found so far.
%
\begin{figure}[tb]
  \includegraphics[width=\columnwidth]{figure/stereo_global.png}
 \caption{Energy plots for the stereo problem. The energy is in the log
 scale. For a multi-threading algorithm, the energy at a given moment is
 defined to be the best solution so far.} \label{fig:stereo_global}
\end{figure}
%
One unusual outcome is that Sequential Alpha Expansion and Parallel
Alpha Expansion (PAE) have the same convergence speed, indicating that
the stereo problem is a very easy one.
%This is well known in a community and this fact was rather expected.
Our approaches with multi-way fusion (SF-SS or SF) are the slowest kind,
because the TRW-S for multi-way fusion is slower than multiple Alpha
Expansion steps, and this stereo problem is too easy to gain benefits
through mulit-way fusion.
%
%Another interesting observation is that the Hierarchical Fusion can
%achieve good energy only when they merge solutions at the top of the
%tree.
%
%  both PAE and SF-MF converge faster
% than sequential method. However, since fusing solutions with multiple
% labels by QPBO is slower than a single $\alpha$-expansion, PAE method
%
% Finally, the line of hierarchy fusion only makes
% jump when fusing on root node of the tree. This is because any fusion
% step on non-root nodes only have partial label information.
%We recoreded both single thread energy and system energy against
%time.
%
Fig. \ref{fig:stereo_threads} shows the energy plot per thread
for SF-MF (ours) and PAE. The graph for PAE again confirms that the
stereo is an easy problem, because all the threads are quickly
converging to a solution before the final sequential fusion.
%shows the per-thread energy
%minimization process in SF-MF and PAE. The per-thread energy in SF-MF
%architecture decreases more uniformly than that of PAE. In the
%scenario where we need to query the best solution so far before the
%whole optimization converges, SF-MF architecture is a better choice.
\begin{figure}[tb]
  \includegraphics[width=\columnwidth]{figure/stereo_threads.png}
  \caption{Energy plots per thread for the stereo proble for SF-MF (left) and
 Parallel Alpha Expansion (right).
} \label{fig:stereo_threads}
\end{figure}

For an easy optimization problem such as stereo with strong unary terms
and submodular pairwise terms, our full architecture with solution
sharing and multiway fusion actually makes convergence slower due to the
overhead of fusion and multi-threading.

% over-sophisticated fusion algorihtm and multi-threading
% overhead. However, we can easily configure the architecture to make it
% better fit the problem, e.g. turn off multiway fusion and/or solution
% sharing.

% We define the energy of
% a parallel optimization system at a certain time as the minimum energy
% of all threads at that time.