### 2016-03-14 paper/stereo.tex
Hang

\section{Swarm Fusion instantiation}
% \hang{I didn't understand the organization of this section at the
%   first glance. Should it be clearer if we put 'Swarm fusion
%   architectures' in front of 'Competing methods' in each problem?}

We compare SF against competing approaches over three problems in
Computer Vision, namely, stereo, optical flow, and layered depthmap
estimation.

% SF framework has high degrees of freedom and the
%challenge is to properly quantify the contributions of various
%algorithmic aspects.

%For this purpose, we have used the three SF architectures illustrated in
%Fig.~\ref{fig:model}.
%% For some problems, we have further limited the capabilities of SF
%% on purpose, to enable effective and fair comparative
%% evaluations. We now explain the details of the three problems.

% We pick three representative problems in Computer Vision to evaluate the
% effectiveness of the Fusion swarm methods against existing methods. We
% control the SF architectures for each problem independently to make
% compa
%
%\yasu{this section is very rough at the moment. requires polishing. but
%algorithms might change as we run experiments.}  We now explain detailed
%implementation of the Swarm Fusion method for three problems in Computer
%Vision (See Fig.~\ref{fig:problem}).  5
\begin{figure}[tb]
  \includegraphics[width=\columnwidth]{figure/problem.pdf} \caption{We
    compare our Swarm Fusion method against competing approaches on a
    depthmap stereo~\cite{middlebury_stereo}, an optical
    flow~\cite{middlebury_optical_flow}, or a layered depthmap
    estimation~\cite{layered_depthmap} problem. In the layered
    depthmap problem, the input is an RGBD image, and the output is
    multiple layeres of depthmaps. Each layer is a piecewise smooth
    parametric surface model.}\label{fig:problem}
\end{figure}

\subsection{Swarm Fusion stereo}
We start with a simple depthmap stereo problem with standard unary and
pairwise terms. We employ submodular pairwise terms to make this
stereo represent relatively ``easy'' MRF inference problems.
%
Unary terms are computed as the average robust photoconsistancy
score~\cite{second_order_stereo} between the reference image and the others
inside a $7\times 7$ pixels window.  Pairwise terms are simple
truncated absolute label difference with maximum label difference
$\sigma_s=4$. The total energy is defined by the sum of the two, while
scaling the pairwise term by a factor of $0.005$. For simplicity we
do not enforce the visibility constraint.

%

\mysubsubsection{Competing methods}

\noindent For simple stereo problems as ours, the sophistication of
photometric consistency function~\cite{mvs_furukawa_survey} makes
unary terms highly informative, thus eliminates the necessity of
adopting sophicticated proposals. Our experiments have also supported
this, where the fusion method (i.e., the use of proposal generation)
rather makes it slow due to the overhead, while not improving final
energy. Therefore, we have chosen algorithms based on Alpha Expansion,
namely single thread Alpha Expansion(AE), Parallel Alpha
Expansion(PAE)~\cite{fusion_moves_for_markov_random_field_optimization}
and Hierarchical Fusion(HF)~\cite{olga_hierarchical_alpha_expansion}
for competing methods. For HF, we use Alpha Expansion when either of
the child node is leaf node (constant label) and QPBO for all other
nodes.

\mysubsubsection{Swarm Fusion architectures}

\noindent The three swarm architectures in Fig.~\ref{fig:model} have
been evaluated: SF-MF (SF without multi-way fusion), SF-SS (SF without
solution sharing), and the standard SF.
%
SF-MF implies $\alpha+\beta=1$, where each thread will expand 4 labels
(repeat $\alpha=1, \beta=0$ for 4 iterations) with Alpha Expansion and
then fuse one solution from other threads ($\alpha=0, \beta=1$) by
QPBO. In the later case, a thread randomly chooses one target thread
to take its solution for fusion. SF-SS implies $\beta=0$, where
$\alpha$ is the free parameter and set to 4. In this case one thread
will fuse 4 labels, together with current solution by TRW-S per
iteration and never exchanges solutions with other threads. We perform
a multi-way fusion at the end to fuse solutions from all the threads
(similar to PAE). For standard SF, we have used ($\alpha=4, \beta=1$).

%
%
%
%
To make the comparison simple, we restrict our proposal generator
in each thread to only generate constant-label proposals within a
subset of all labels.
%
% Due to the lack of multi-way fusion, the swarm fusion architecture for
% the stereo problem is represented by the middle example at the bottom of
% Fig.~\ref{fig:model}.