### 2016-03-14 paper/conclusion.tex
spell checking

\section{Conclusion and future directions}
We have proposed a novel MRF inference framework, Swarm Fusion, in
parallel computing environments. The framework is general and makes
popular inference techniques such as Alpha Expansion, Fusion Move,
Parallel Alpha Expansion, and Hierarchical Fusion, its special cases. Our
experiments have revealed that the framework exploits parallel
computational resources and achieves faster convergence, especially for
challenging problems.  Our first future work is to conduct experiments
on cloud computing environments, in particular, the MapReduce
programming model, where the roles of mappers and reducers exactly
correspond to the processes of parallel multi-way fusion and solution
sharing, respectively.  Another future work is the automatic
configuration of the Swarm Fusion architecture.  Our experiments have
shown that optimal architectures are different for different problems.
An interesting direction is to adaptively change its architecture during
the computation, for example, switching to simple parallel
alpha-expansion for easy problems, or increasing the rate of solution
exchanges when solutions vary significantly across threads.
%
Parallel MRF inference has been a relatively under-explored topic in
Computer Vision. Our idea is very simple, where adapting the Swarm Fusion
framework require minimal coding. We believe that this paper would
immediately benefit tens of thousands of Computer Vision researchers or
engineers in the world, who currently solve MRF problems. We will share
our source code with the community.