### 2016-03-15 paper/related_work.tex
Hang

\section{Related work}

MRF inference has been a very active field in Computer Vision with
extensive literature. We refer the readers to survey articles for
comprehensive
reviews~\cite{middlebury_mrf,comparative_study_of_modern_inference}, and
here focus our description on closely related topics.

\mysubsubsection{Parallel Alpha-Expansion}

\noindent Lempitsky et
al.~\cite{fusion_moves_for_markov_random_field_optimization}
introduces parallel computation to the alpha-expansion technique,
where multiple threads simultaneously fuse mutually exclusive sets of
labels. Kumar et al.~\cite{hierarchical_graph_cuts_kumar_and_koller},
Delong et al.~\cite{delong_hierarchical_fusion}, and Veksler et
al.~\cite{olga_hierarchical_alpha_expansion} investigated hierarchical
approachs, where labels can be simultaneously fused from the bottom to
the top in a tree of labels.
% 
Instead of taking a hierarchical approach, Batra et
al.~\cite{Dhruv_pushmeet_making_the_right_move} adaptively computed an
effective sequence of labels to explore. This technique can be combined
with parallel alpha-expansion techniques to further obtain speed-up.
%
Strictly speaking, these approaches are not in the family of Fusion Move
methods (FM), because they only consider constant label proposals. Our
approach is a generalization of FM.

\mysubsubsection{Parallel MAP inference}

\noindent The core MAP inference itself can be parallelized.  Strandmark
et al.~\cite{strandmark_parallel} parallelized graph-cuts.
%
Message passing algorithms are friendly to GPU implementation and can
exploit the power of parallel computation.
%
%
While state-of-the-art optimization libraries are often freely available
for non-commercial purposes, most companies have to develop and maintain
in-house implementation of these algorithms.  The core optimization
libraries are very complex and their modifications require significant
engineering investments. In contrast, our idea is extremely simple and
easily reproducible by standard engineers.

% \footnote{GPU speeds-up message-passing algorithms via parallel
% computation. However, these algorithms need to store all the messages
% and states and cannot handle problems with a large label
% space~\cite{layered_depthmap}.}

\mysubsubsection{Fusion Move methods}

\noindent FM was first introduced by Lempitsky et al.~\cite{fusion_flow} in solving the
optical flow problem. FM has been effectively used to solve challenging
problems in Computer Vision such as stereo with second order smoothness
priors~\cite{second_order_stereo}, stereo with parameteric surface fitting and
segmentation (i.e. Surface Stereo)~\cite{surface_stereo}, and multicut
partitioning~\cite{fusion_moves_for_correlation_clustering}.
%
FM has two main advantages over other general inference
techniques~\cite{TRW-S,loopy_belief_propagation}. First, FM allows us to
exploit domain-specific knowledge by customizing proposal generation
schemes. Second, FM can handle problems with very large label spaces
(and even real-values variables), because the core optimization
solves a sequence of binary decision problems.
%
In contrast, methods like message passing algorithms need to maintain
messages and beliefs for the entire label space all the time.
%
Although conceptually straightforward, we are not aware of {\it Parallel
  Fusion Move (PFM)} algorithms that fuse solution proposals, as opposed to
labels, in parallel. This paper seeks to fully unleash the power of
parallel computation based on FM in the most general setting.
% for the most general FM via a more general framework.

\mysubsubsection{Evolutionary algorithms and Particle Swarm Optimization}

\noindent
Genetic algorithms (GA)~\cite{ga} and Particle Swarm
Optimization (PSO)~\cite{pso} maintain multiple solutions and improve
them over time.
%
GA or PSO has been used to produce great empirical results,
e.g. in hand tracking~\cite{pushmeet_hand_tracking}.
%
At high level, our strategy is similar in spirit. However, GA or PSO
rather arbitrarily copies parts of the solutions or makes random movements
in each step (i.e., limited theoretical justification).
%
Our approach directly optimizes the objective function to improve
solutions.