### 2016-03-15 paper/result_layered_depthmap.tex
minor changes

\mysubsubsection{Layered depthmap estimation}

\noindent We have used ``ours\_1'' data in~\cite{layered_depthmap} for
the experiments. Fig.~\ref{fig:layered_depthmap_convergence} shows
that Fusion Move, Parallel Fusion Move and SF-MF all got stuck in local
minima, which is due to the lack of multi-way fusion.  Layered depthmap
estimation is a challenging problem with very large solution space. The
binary fusion of solution proposals is too restrictive to make any
improvements.  This coincides with the observation in
\cite{layered_depthmap} that binary fusion of proposal solutions is not
as powerful as their subspace fusion which is a special form of
multi-way fusion here. Lastly, solution sharing also plays an important
role for this challenging problem for exactly the same reason, as SF
performs much better than SF-SS.

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figure/layered_depthmap_convergence.png}
  \caption{Energy plots for the layered depthmap estimation
    problem. Both the multi-way fusion and the solution sharing are important
    for this challenging problem.}\label{fig:layered_depthmap_convergence}
\end{figure}

\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{0.49\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{figure/layered_depthmap_by_alpha.png}
    \caption{}
    \label{fig:layered_depthmap_by_alpha}
  \end{subfigure}  
  \begin{subfigure}[b]{0.49\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{figure/layered_depthmap_by_N.png}
    \caption{}
    \label{fig:layered_depthmap_by_N}
  \end{subfigure}
  \caption{(a) Energy plots for layered depthmap estimation with varying
    $\alpha$. (b) Energy plots for layered depthmap estimation with varying number of threads N.}
  \label{fig:layered_depthmap_by_alpha_N}
\end{figure}

%
% From the plot, we can see that, Fusion Move, Parallel Fusion Move, and
%SF-MF all stalk at a high energy state.
To further study the effects of multi-way fusion, we have varied the
value of $\alpha$ which controls the number of solution proposals to be
fused in SF-SS model (See
Fig.~\ref{fig:layered_depthmap_by_alpha}). Note that we have used SF-SS
instead of SF to disable solution sharing and better observe the effects
of multi-way fusion.
% SF-SS model (disable solution sharing to
%better observe the effect of multi-way fusion) while keeping other
%parameters the same and plot the energy minimization process in figure
It is interesting to see that more multi-way fusion takes longer to
converge, but finds a lower energy state at the end.

At the end we examined the role of multi-threading by varying the number of threads N in our most general model SF (See Fig.~\ref{fig:layered_depthmap_by_N})~\footnote{While keeping other parameters the same, we have to change $\beta$ with $N$ because of the constraint $\beta \leq N-1$. We always use $\beta = N-1$ in this experiment.}. The plot shows that SF finds low energy faster with more threads. In Fusion Move like methods, parallelism speeds up convergence by exploring solution proposals faster. Due to the inherited randomness in the proposal generation scheme, the speed up is often not proportional to the number of threads.
%We can see that as the number of
%ways to fuse becomes larger, each fusion step takes longer time, but the
%chance of finding a lower energy state increases.