### 2016-03-15 paper/stereo.tex

\section{Swarm Fusion instantiation}
% \hang{I didn't understand the organization of this section at the
%   first glance. Should it be clearer if we put 'Swarm fusion
%   architectures' in front of 'Competing methods' in each problem?}

We compare SF against competing approaches over three problems in
Computer Vision, namely, stereo, optical flow, and layered depthmap
estimation.

% SF framework has high degrees of freedom and the
%challenge is to properly quantify the contributions of various
%algorithmic aspects.

%For this purpose, we have used the three SF architectures illustrated in
%Fig.~\ref{fig:model}.
%% For some problems, we have further limited the capabilities of SF
%% on purpose, to enable effective and fair comparative
%% evaluations. We now explain the details of the three problems.

% We pick three representative problems in Computer Vision to evaluate the
% effectiveness of the Fusion swarm methods against existing methods. We
% control the SF architectures for each problem independently to make
% compa
%
%\yasu{this section is very rough at the moment. requires polishing. but
%algorithms might change as we run experiments.}  We now explain detailed
%implementation of the Swarm Fusion method for three problems in Computer
%Vision (See Fig.~\ref{fig:problem}).  5
\begin{figure}[tb]
  \includegraphics[width=\columnwidth]{figure/problem.pdf} \caption{We
    compare our Swarm Fusion method against competing approaches on the
    depthmap stereo~\cite{middlebury_stereo}, the optical
    flow~\cite{middlebury_optical_flow} and the layered depthmap
    estimation~\cite{layered_depthmap} problem. In the layered
    depthmap problem, the input is an RGBD image, and the output is
    multiple layers of depthmaps. Each layer is a piecewise smooth
    parametric surface model.}\label{fig:problem}
\end{figure}

\subsection{Swarm Fusion stereo}
We start with a simple depthmap stereo problem with standard unary and
pairwise terms. We employ submodular pairwise terms to make this
stereo represent relatively ``easy'' MRF inference problems.
%
Unary terms are computed as the average robust photoconsistancy
score~\cite{second_order_stereo} between the reference image and the others
inside a $7\times 7$ pixels window.  Pairwise terms are simple
truncated absolute label difference with maximum label difference
$\sigma_s=4$. The total energy is defined by the sum of the two, while
scaling the pairwise term by a factor of $0.005$. For simplicity we
do not enforce the visibility constraint.

%

\mysubsubsection{Competing methods}

\noindent For simple stereo problems with submodular energy as ours,
the sophistication of photometric consistency
function~\cite{mvs_furukawa_survey} makes unary terms highly
informative, where highly efficient algorithms such as graph-cuts
exist. Our experiments have also supported this, where the fusion
method (i.e., the use of proposal generation) rather makes it slow due
to the overhead, while not improving final energy. Therefore, we have
chosen algorithms based on Alpha-Expansion, namely single thread Alpha
Expansion(AE), Parallel Alpha
Expansion(PAE)~\cite{fusion_moves_for_markov_random_field_optimization}
and Hierarchical Fusion(HF)~\cite{olga_hierarchical_alpha_expansion}
to be competing methods. For HF, we use Alpha-Expansion when either of
the child node is leaf node (constant label) and QPBO for all other
nodes.

\mysubsubsection{Swarm Fusion architectures}

\noindent The three swarm architectures in Fig.~\ref{fig:model} have
been evaluated: SF-MF (SF without multi-way fusion), SF-SS (SF without
solution sharing), and the standard SF.
%
SF-MF implies $\alpha+\beta=1$, where each thread will expand 4 labels
(repeat $\alpha=1, \beta=0$ for 4 iterations) with Alpha Expansion and
then fuse one solution from other threads ($\alpha=0, \beta=1$) by
QPBO. In the later case, a thread randomly chooses one solution from
the solution pool for fusion. SF-SS implies $\beta=0$, where $\alpha$
is the free parameter and set to 4. In this case one thread will fuse
4 labels, together with current solution by TRW-S per iteration and
never exchanges solutions with other threads. We perform a multi-way
fusion at the end to fuse solutions from all the threads (similar to
PAE). For standard SF, we have used ($\alpha=4, \beta=1$).

%
%
%
%
To make the comparison simple, we restrict our proposal generator
in each thread to only generate constant-label proposals within a
subset of all labels.
%
% Due to the lack of multi-way fusion, the swarm fusion architecture for
% the stereo problem is represented by the middle example at the bottom of
% Fig.~\ref{fig:model}.